Simulation study in rectangular domain

### 2024-04-15 R/simulation_study/simulation_study_rect.Rmd
Commit Latex and simulation study changes

---
title: "Simulation study in rectangular domain"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

@@

#used in function split_term to get covariates names from the formulas
library(stringr)

@@

@@

#number of tracks
n_samples=1

@@

#generate random initial points
x0=matrix(rep(NA,n_samples*2),ncol=2)
colnames(x0)=c("x1","x2")

for (i in 1:n_samples) {
  #choose location uniformly in the domain
  x=c(runif(1,min=1.5,max=8.5),runif(1,min=1.5,max=8.5))
  x0[i,]=x
  }
```

@@

```{r}
#observation times
Tmax=3*24 #max time in hour

@@

```{r}
#no measurement error (we cannot put 0 because only the log of this quantity is used)
sigma_obs_0=0.005
H0_hf=array(rep(sigma_obs_0^2*diag(2),n_hf),dim=c(2,2,n_hf))
#GPS measurement error 
sigma_obs_1=0.03
H1_hf=array(rep(sigma_obs_1^2*diag(2),n_hf),dim=c(2,2,n_hf))

@@

#constant tau
ftau_constant=function(cov_data,tau=1) {
  return (tau)
}

@@

nu=4
```

@@

```{r}
sp_df=5
#number of points
n <- 500

@@

#fit with bivariate splines te
m1 <- gam(y~te(theta,DistanceShore,k=sp_df))

@@

#fit with bivariate splines te
m2 <- gam(y~te(theta,ExpShore,k=sp_df))

@@

```
### Get estimated spline coefficients

@@

### Generate sample
```{r, cache=TRUE,results='hide'}

#initialize data frame
data_rect_hf_ne=data.frame("y1"=NA,"y2"=NA,"time"=NA,"p1"=NA,"p2"=NA,"ID"=NA)

for (i in 1:n_samples) {

  res=sim_theta_CRCVM(ftau_constant,fomega,nu,log_sigma_obs=log(sigma_obs_0),v0,x0[i,],times_hf,land=domain,verbose=FALSE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_shore=res$shore[,c("p1","p2")]
  data_rect_hf_ne=rbind(data_rect_hf_ne,cbind(data_sim,data_shore))
}

#for the first value we cannot compute the angle theta
data_rect_hf_ne=data_rect_hf_ne[-1,]

@@

```{r}
count=0
for (id in unique(data_rect_hf_ne$ID)){
  sub_data=data_rect_hf_ne[data_rect_hf_ne$ID==id,]
  if (nrow(sub_data) < n_hf) {
    print(id)
    count=count+1
  }
}
cat(count/n_samples*100,"percent of the samples reached land")
```

@@

plot_samples_rect_hf_ne=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_hf_ne[data_rect_hf_ne$ID==1,],mapping=aes(y1,y2),size=0.1)+
  geom_path(data=data_rect_hf_ne[data_rect_hf_ne$ID==1,],mapping=aes(y1,y2),size=0.1)+
  geom_point(data = data_rect_hf_ne[data_rect_hf_ne$ID==1,]%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

@@

#initialize data frame
data_rect_hf_e=data.frame("y1"=NA,"y2"=NA,"time"=NA,"p1"=NA,"p2"=NA,"ID"=NA)

for (i in 1:n_samples) {

  res=sim_theta_CRCVM(ftau_constant,fomega,nu,log_sigma_obs=log(sigma_obs_1),v0,x0[i,],times_hf,land=domain,verbose=FALSE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_shore=res$shore[,c("p1","p2")]
  data_rect_hf_e=rbind(data_rect_hf_e,cbind(data_sim,data_shore))
}

#for the first value we cannot compute the angle theta
data_rect_hf_e=data_rect_hf_e[-1,]

@@

```{r}
count=0
for (id in unique(data_rect_hf_ne$ID)){
  sub_data=data_rect_hf_e[data_rect_hf_e$ID==id,]
  if (nrow(sub_data) < n_hf) {
    print(id)
    count=count+1
  }
}
cat(count/n_samples*100,"percent of the samples reached land")
```

@@

plot_samples_rect_hf_e=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_hf_e[data_rect_hf_e$ID==1,],mapping=aes(y1,y2),size=0.1)+
  geom_path(data=data_rect_hf_e[data_rect_hf_e$ID==1,],mapping=aes(y1,y2),size=0.1)+
  geom_point(data = data_rect_hf_e[data_rect_hf_e$ID==1,]%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
    xlab("x") + ylab("y")

@@

add_covs=function(data) {
  #data has at least columns "time","y1" "y2", "p1","p2"
  #this function modifies data to add columns "theta", "DistanceShore" and "ExpShore"

@@

data_rect_hf_ne=add_covs(data_rect_hf_ne)

data_rect_hf_e=add_covs(data_rect_hf_e)
```

## Estimate from high frequency data without low error

@@

#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,ExpShore,k=sp_df,bs="cs"))
par0 <- c(0,0,1,4,0)

crcvm_rect_hf_ne1 <- SDE$new(formulas = formulas,data = data_rect_hf_ne[data_rect_hf_ne$ID==1,],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2","nu","tau"),
                      other_data=list("H"=H0_hf))
crcvm_rect_hf_ne1$fit(method="BFGS")
```

@@

```{r,cache=TRUE}
#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,DistanceShore,k=sp_df,bs="cs"))
par0 <- c(0,0,1,4,0)

>>>>

---
title: "Simulation study in rectangular domain"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
  date: "2024-03-13"
---

@@

@@

# global variables

```{r}
#number of tracks
N_BATCH=1
N_ID=3

#degree of freedom in tensor splines
SP_DF=5
```

@@

@@

#generate random initial points
x0=matrix(rep(NA,N_ID*2),ncol=2)
colnames(x0)=c("x1","x2")

for (i in 1:N_ID) {
  #choose location uniformly in the domain
  x=c(runif(1,min=1.5,max=8.5),runif(1,min=1.5,max=8.5))
  x0[i,]=x
}

@@

```{r}
#observation times
Tmax=1*24 #max time in hour

@@

```{r}
#no measurement error (we cannot put 0 because only the log of this quantity is used)
sigma_obs_0=0.005
H0_hf=array(rep(sigma_obs_0^2*diag(2),n_hf*N_ID),dim=c(2,2,n_hf*N_ID))
#GPS measurement error 
sigma_obs_1=0.03
H1_hf=array(rep(sigma_obs_1^2*diag(2),n_hf*N_ID),dim=c(2,2,n_hf*N_ID))

@@

#constant tau
ftau_constant=function(cov_data,tau=1,sigma_re=0.1) {
  b=rnorm(1,mean=0,sd=sigma_re)
  return (tau+b)
}

@@

#constant nu
fnu_constant=function(cov_data,nu=1,sigma_re=0.3) {
    b=rnorm(1,mean=0,sd=sigma_re)
  return (nu+b)
}

@@

```{r}
SP_DF=5
#number of points
n <- 500

@@

#fit with bivariate splines te
m1 <- gam(y~te(theta,DistanceShore,k=SP_DF))

@@

#fit with bivariate splines te
m2 <- gam(y~te(theta,ExpShore,k=SP_DF))

@@

### Get estimated spline coefficients

@@

### Generate sample

#initialize data frame
data_rect_hf_ne=list()

for (k in 1:N_BATCH) {

  data=data.frame("y1"=NA,"y2"=NA,"time"=NA,"p1"=NA,"p2"=NA,"ID"=NA)

  for (i in 1:N_ID) {

    res=sim_theta_CRCVM(ftau_constant,fomega,fnu_constant,log_sigma_obs=log(sigma_obs_0),v0,x0[i,],times_hf,land=domain,verbose=FALSE)

    data_sim=res$sim
    data_sim$ID=factor(rep(i,length(data_sim$y1)))
    data_shore=res$shore[,c("p1","p2")]
    data=rbind(data,cbind(data_sim,data_shore))
  }

  #for the first value we cannot compute the angle theta
  data=data[-1,]
  data_rect_hf_ne[[paste("BATCH",k,sep="")]]=data
}

@@

```{r}
count=0
for (k in 1:length(names(data_rect_hf_ne))) {

  data=data_rect_hf_ne[[paste("BATCH",k,sep="")]]
  for (id in unique(data$ID)) {
    sub_data=data[data$ID==id,]
    if (nrow(sub_data) < n_hf) {
      count=count+1
      cat("BATCH",k,id,"\n",sep="")
    }
  }
}
cat(count/(N_BATCH*N_ID)*100,"percent of the samples reached land")
```

### plot one batch

@@

plot_samples_rect_hf_ne=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_path(data=data_rect_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_point(data = data_rect_hf_ne$BATCH1%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

@@

#initialize data frame
data_rect_hf_e=list()

for (k in 1:N_BATCH) {

  data=data.frame("y1"=NA,"y2"=NA,"time"=NA,"p1"=NA,"p2"=NA,"ID"=NA)

  for (i in 1:N_ID) {

    res=sim_theta_CRCVM(ftau_constant,fomega,fnu_constant,log_sigma_obs=log(sigma_obs_1),v0,x0[i,],times_hf,land=domain,verbose=FALSE)

    data_sim=res$sim
    data_sim$ID=factor(rep(i,length(data_sim$y1)))
    data_shore=res$shore[,c("p1","p2")]
    data=rbind(data,cbind(data_sim,data_shore))
  }

  #for the first value we cannot compute the angle theta
  data=data[-1,]
  data_rect_hf_e[[paste("BATCH",k,sep="")]]=data
}
```

@@

```{r}
count=0
for (k in 1:length(names(data_rect_hf_e))) {

  data=data_rect_hf_e[[paste("BATCH",k,sep="")]]
  for (id in unique(data$ID)) {
    sub_data=data[data$ID==id,]
    if (nrow(sub_data) < n_hf) {
      count=count+1
      cat("BATCH",k,id,"\n",sep="")
    }
  }
}
cat(count/(N_BATCH*N_ID)*100,"percent of the samples reached land")
```

@@

plot_samples_rect_hf_e=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_hf_e$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_path(data=data_rect_hf_e$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_point(data = data_rect_hf_e$BATCH1%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

@@

add_covs=function(data) {
  #data has at least columns "time","y1" "y2", "p1","p2"
  #this function new data frame with three more columns "theta", "DistanceShore" and "ExpShore"

@@

for (k in 1:N_BATCH) {

  #dataframes for this batch
  df_ne=data_rect_hf_ne[[paste("BATCH",k,sep="")]]
  df_e=data_rect_hf_e[[paste("BATCH",k,sep="")]]

  #add covariates columns
  df_ne=add_covs(df_ne)
  df_e=add_covs(df_e)

  #change in the list of batch
  data_rect_hf_ne[[paste("BATCH",k,sep="")]]=df_ne
  data_rect_hf_e[[paste("BATCH",k,sep="")]]=df_e
}

## Estimate from high frequency data with low error

@@

#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,ExpShore,k=SP_DF,bs="cs"))
par0 <- c(0,0,1,4,0)

crcvm_rect_hf_ne1 <- SDE$new(formulas = formulas,data = data_rect_hf_ne$BATCH1,type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H0_hf))
crcvm_rect_hf_ne1$fit(method="BFGS")
```

@@

```{r,cache=TRUE}
#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,DistanceShore,k=SP_DF,bs="cs"))
par0 <- c(0,0,1,4,0)
